Cloud Supercomputing

### 2019-03-25 writing/paper/paper.tex
Proof reading.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
\section{Introduction}
\newcommand{\diffCoeff}{\varepsilon}
\newcommand{\hyperFluxDef}{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure) \\
    \Qj \QZZ
  \end{pmatrix}
}
\newcommand{\viscFluxDef}{
  \begin{pmatrix}
    0\\
     \stressT (\Q, \gradQ)  \\
     \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}\\
     0\\
   \end{pmatrix}
}%
We solve the reactive compressible Navier-Stokes equations in the conservative form% \pcref{eq:conservation-law-gradient}
\begin{equation}
 \label{eq:equation-set} 
%  \begin{array}{l}
%  \text{mass cons.} \\
%  \text{momentum cons.} \\
%  \text{energy cons.} \\
%  \text{cont.\ gas} 
% \end{array}
% :
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE\\
    \QZ
    \end{pmatrix}}_{\Q}
  +
  \divergence{
  \underbrace{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}_{\flux(\Q, \gradQ)}}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    \source[\Qj]\\
    \source[\QE]\\
    \source[\QZ]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}
with vector of conserved quantitites $\Q$, flux $\flux(\Q, \gradQ)$ and source $\source(\Q)$.
The conserved quantities
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right),
\end{equation}
are the density $\Qrho$, the two or three-dimensional momentum $\Qj$, the energy density $\QE$ and the mass fraction of the chemical reactant $\QZ$.
The pressure $\pressure$ is given by the equation of state of an ideal reacting gas
\begin{equation}
  \label{eq:eos}
  \pressure = (\gamma - 1) \left(\QE - \frac{1}{2} \left(\Qv \cdot \Qj \right)  - q_0 \QZ - gz \right).
\end{equation}
and relates to the temperature $T$ by the thermal equation of state
\begin{equation}
  \label{eq:temperature}
  \pressure = \Qrho R T,
\end{equation}
where $R$ is the specific gas constant of a fluid.

The term $q_0 \QZ$ is the chemical energy with a heat release $q_0$~\cite{helzel2000modified}, the term $gz$ is the geopotential height with the gravity of Earth $g$~\cite{giraldo2008study}.
We set $\QZZ$ and $q_0$ to zero for scenarios without a reactive component.
We model the diffusive effects of the flow by the standard stress tensor
\begin{equation}
  \label{eq:stress-tensor}
  \stressT(Q, \nabla Q) =
  \mu
  \left(
  \left(\nicefrac{2}{3} \divergence{\Qv} \right) -
  \left( \gradient{\Qv} + \gradient{\Qv}^\intercal \right)
  \right),
\end{equation}
with viscosity $\mu$.

Many realistic, atmospheric flows can be described by a pertubation over a background state that is in hydrostatic equilibrium
\newcommand{\backgroundPressure}{\overline{\pressure}}
\newcommand{\backgroundRho}{\overline{\Qrho}}
\begin{equation}
  \label{eq:hydrostatic-balance}
  \pdv{}{z} \backgroundPressure{\left (z \right )} = -g \backgroundRho(z),
\end{equation}
i.e.\ a state where the pressure differences are exactly balanced with the gravitational source term.
The momentum equation is dominated by the background flow in this case.
This leads to numerical instabilities.
To avoid this, we split the pressure $\pressure = \backgroundPressure + \pressure'$ and density $\Qrho = \backgroundRho + \Qrho'$, arriving at the modified equation
\begin{equation}
  \label{eq:momentum-equation-split}
  \pdv{\Qj}{t}+ \divergence{ \left(
    \Qv \otimes \Qj + \bm{I} \pressure'
    \right)
  } + \viscFlux_\Qj
  =
  -g \bm{k} \Qrho'.
\end{equation}

@@

In the rest of the paper, we make the following contributions:
\begin{itemize}
\item Efficient, explicit \aderdg{}-scheme with shared and distributed memory parallelization.
\item An \textsc{amr}-criterion that is based on the detection of outlier cells w.r.t.\ their total variation.
\item Evaluation with standard \textsc{cfd} scenarios, atmospheric and reactive flows.
\end{itemize}

\section{Methods}
\todo{Write \aderdg{}/\exahype{} section!}
Maybe numerics (dg+muscl?), at least ref.
Exahype?!

@@

We use the Arnold-Beltrami-Childress (\textsc{abc}) flow as a simple three-dimensional flow~\cite{tavelli2016staggered}.
It also has an exact solution under the assumption of incompressibility
\begin{align}
  \label{eq:abc-flow}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \phantom{-} \exp(-1\mu t)
  \begin{pmatrix}
    \sin(z) + \cos(y)\\
    \sin(x) + \cos(z)\\
    \sin(y) + \cos(x)
  \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= -\exp(-2 \mu t) \, \left(\cos(x)\sin(y) + \sin(x)\cos(z) + \sin(z)\cos(y)\right)
  + C
  \end{split}
\end{align}
Here, $C = \nicefrac{100}{\gamma}$ is an arbitrary constant.
We use a viscosity of $\mu = 0.01$ and Cauchy boundary conditions.

As a final example of standard flow scenarios, we consider the lid-driven cavity flow which initial conditions are given by a fluid at rest, with $\Qrho = 1, \pressure(\bm{x}) = \nicefrac{100}{\gamma}$.

@@

Our atmospheric flow scenarios are described in terms of the potential energy
\begin{equation}
  \potT = T \left( \frac{p_0}{p} \right)^{R/c_p},
\end{equation}
with reference pressure $p_o = \SI{10e5}{\Pa}$~\cite{muller2010adaptive,giraldo2008study}.
For all remaining scenarions, we use the constants
\begin{equation}\label{eq:atmosphere-constants}
    \gamma = 1.4 ,\qquad \Pr =  0.71 ,\qquad R = 287.058 ,\qquad p_0 = 10000, \qquad g = 9.8.
\end{equation}

We compute the initial background density and pressure by inserting the assumption of a constant background energy in \cref{eq:hydrostatic-balance}.
The background atmosphere is then perturbed.
\todo{What about our weird boundary condition for heat flux and energy?}
Let
\begin{equation}
  \label{eq:radius}
  r^2 = (x - x_0)^2 + (z - z_0)^2
\end{equation}
denote the difference between spatial positions $x,z$ and the center of a bubble $x_c, z_c$.
For our first scenario, the cosine bubble, we use a pertubation of the form
\begin{equation}
  \label{eq:cos-pertubation-alt}
  \pertubationPotT = 
    \nicefrac{A}{2} \left( 1 + \cos(\pi r) \right) \left[ r \leq a \right]
\end{equation}
\begin{align}
  \label{eq:cos-pertubation}
  \pertubationPotT &= \begin{cases}
    \nicefrac{A}{2} \left[ 1 + \cos(\pi r) \right] & r \leq a, \\
    0 & r > a,
    \end{cases}
\end{align}
where $A$ denotes the maximal perturbation and $a$ is the size of the bubble.
We use the constants
\begin{equation}\label{eq:cosine-bubble}
  A = 0.5 \quad a = \SI{250}{\m} \quad x_c = \SI{500}{\m} \quad z_c = \SI{350}{\m}.
\end{equation}
We use a constant viscosity of $\mu = 0.1$.
For our second scenario, the cold bubble on top of a warm bubble, we use the pertubations
\begin{equation}
  \label{eq:bubbles-pertubation}
  \pertubationPotT =
  \begin{cases}
    A & r \leq a, \\
    A \exp \left( - \frac{(r-a)^2}{s^2} \right) & r > a,
    \end{cases}
\end{equation}
where $s$ is the decay rate and $r$  is the radius to the center~\pcref{eq:radius}.
We have two bubbles, with constants
\begin{equation}
  \label{eq:bubbles-values}
\begin{alignedat}{6}
  & \text{warm:} \qquad && A = 0.5, \quad&& a = \SI{150}{\m}, \quad&& s = \SI{50}{\m}, \quad&& x_c = \SI{500}{\m,} \quad&& z_c = \SI{300}{\m},\\
  & \text{cold:} \qquad && A = -0.15, \quad&& a = \SI{0}{\m}, \quad&& s = \SI{50}{\m}, \quad&& x_c = \SI{560}{\m}, \quad&& z_c = \SI{640}{\m}.
  \end{alignedat}
\end{equation}
Similar to~\cite{muller2010adaptive} and to the previous scenario, we use a constant viscosity of $\mu = 0.1$ to regularize the solution.
To ensure that the atmosphere stays in hydrostatic balance~\cite{giraldo2008study}, we need to impose the viscous heat flux\todo{Show full bc?}
\begin{equation}
  \label{eq:atmosphere-bc}
  \viscFlux_{\QE} = \kappa \pdv{\overline{T}}{z}% =
%\kappa \frac{R \overline{\theta} \left(\frac{\overline{p}{\left (z \right )}}{p_{0}}\right)^{\frac{R}{c_{p}}} \frac{d}{d z} \overline{p}{\left (z \right )}}{c_{p} \overline{p}{\left (z \right )}}.
\end{equation}

>>>>

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
\section{Introduction}
\newcommand{\diffCoeff}{\varepsilon}
\newcommand{\hyperFluxDef}{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure) \\
    \Qj \QZZ
  \end{pmatrix}
}
\newcommand{\viscFluxDef}{
  \begin{pmatrix}
    0\\
     \stressT (\Q, \gradQ)  \\
     \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}\\
     0\\
   \end{pmatrix}
}%
We solve the reactive compressible Navier-Stokes equations in the conservative form% \pcref{eq:conservation-law-gradient}
\begin{equation}
 \label{eq:equation-set} 
%  \begin{array}{l}
%  \text{mass cons.} \\
%  \text{momentum cons.} \\
%  \text{energy cons.} \\
%  \text{cont.\ gas} 
% \end{array}
% :
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE\\
    \QZ
    \end{pmatrix}}_{\Q}
  +
  \divergence{
  \underbrace{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}_{\flux(\Q, \gradQ)}}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    \source[\Qj]\\
    \source[\QE]\\
    \source[\QZ]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}
with vector of conserved quantitites $\Q$, flux $\flux(\Q, \gradQ)$ and source $\source(\Q)$.
The conserved quantities
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right),
\end{equation}
are the density $\Qrho$, the two or three-dimensional momentum $\Qj$, the energy density $\QE$ and the mass fraction of the chemical reactant $\QZ$.
In \cref{eq:equation-set}, the rows are the conservation of mass, the conservation of momentum, the conservation of energy and the continuity equation of the reactant.

The pressure $\pressure$ is given by the equation of state of an ideal reacting gas\todo{Describe gravitation?}
\begin{equation}
  \label{eq:eos}
  \pressure = (\gamma - 1) \left(\QE - \frac{1}{2} \left(\Qv \cdot \Qj \right)  - q_0 \QZ - gz \right).
\end{equation}
The term $q_0 \QZ$ is the chemical energy with a heat release $q_0$~\cite{helzel2000modified}, the term $gz$ is the geopotential height with the gravity of Earth $g$~\cite{giraldo2008study}.
We set $\QZZ$ or $q_0$ to zero for scenarios without a reactive component.
The temperature $T$ relates to the pressure by the thermal equation of state
\begin{equation}
  \label{eq:temperature}
  \pressure = \Qrho R T,
\end{equation}
where $R$ is the specific gas constant of a fluid.

We model the diffusivity of the flow by the stress tensor
\begin{equation}
  \label{eq:stress-tensor}
  \stressT(Q, \nabla Q) =
  \mu
  \left(
  \left(\nicefrac{2}{3} \divergence{\Qv} \right) -
  \left( \gradient{\Qv} + \gradient{\Qv}^\intercal \right)
  \right),
\end{equation}
with constant viscosity $\mu$.
The heat diffusion is governed by the coefficient
\begin{equation}
  \label{eq:heat-conduction-coeff}
  \kappa = \frac{\mu \gamma}{\Pr} \frac{1}{\gamma - 1} R = \frac{\mu c_p}{\Pr},
\end{equation}
where the ratio of specific heats $\gamma$, the heat capacity at constant pressure $c_p$ and the Prandtl number $\Pr$ depend on the fluid.

Many realistic atmospheric flows can be described by a pertubation over a background state that is in hydrostatic equilibrium
\newcommand{\backgroundPressure}{\overline{\pressure}}
\newcommand{\backgroundRho}{\overline{\Qrho}}
\begin{equation}
  \label{eq:hydrostatic-balance}
  \pdv{}{z} \backgroundPressure{\left (z \right )} = -g \backgroundRho(z),
\end{equation}
i.e.\ a state where the pressure gradient is exactly balanced with the gravitational source term.
The momentum equation is dominated by the background flow in this case.
This leads to numerical instabilities.
To avoid this, we split the pressure $\pressure = \backgroundPressure + \pressure'$ into a sum of the background pressure $\backgroundPressure(z)$ and pertubation $\pressure'(\bm{x}, t)$.
We split the density $\Qrho = \backgroundRho + \Qrho'$ in the same manner and arrive at
\begin{equation}
  \label{eq:momentum-equation-split}
  \pdv{\Qj}{t}+ \divergence{ \left(
    \Qv \otimes \Qj + \bm{I} \pressure'
    \right)
  } + \viscFlux_\Qj
  =
  -g \bm{k} \Qrho'.
\end{equation}

@@

In the rest of the paper, we make the following contributions:
\begin{itemize}
\item We present an efficient, explicit \aderdg{}-scheme with shared and distributed memory parallelization for our equation set.
\item We show an \textsc{amr}-criterion that is based on the detection of outlier cells w.r.t.\ their total variation.
\item We evaluate our implementation with standard \textsc{cfd} scenarios, atmospheric and reactive flows.
\end{itemize}

\section{Methods}
\todo{Write \aderdg{}/\exahype{} section! Riemann solver + timestep}
Maybe numerics (dg+muscl?), at least ref.
Exahype?!

@@

We use the Arnold-Beltrami-Childress (\textsc{abc}) flow as a simple three-dimensional flow~\cite{tavelli2016staggered}.
It also has an exact solution under the assumption of incompressibility
\begin{align}
  \label{eq:abc-flow}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \phantom{-} \exp(-1\mu t)
  \begin{pmatrix}
    \sin(z) + \cos(y)\\
    \sin(x) + \cos(z)\\
    \sin(y) + \cos(x)
  \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= -\exp(-2 \mu t) \, \left(\cos(x)\sin(y) + \sin(x)\cos(z) + \sin(z)\cos(y)\right)
  + C.
  \end{split}
\end{align}
Here, $C = \nicefrac{100}{\gamma}$ is an arbitrary constant.
We use a viscosity of $\mu = 0.01$ and Cauchy boundary conditions.

As a final example of standard flow scenarios, we consider the lid-driven cavity flow where the fluid is at rest initially, with $\Qrho = 1$ and $ \pressure(\bm{x}) = \nicefrac{100}{\gamma}$.

@@

Our atmospheric flow scenarios are described in terms of the potential energy
\begin{equation}
  \potT = T \left( \frac{p_0}{p} \right)^{R/c_p},
\end{equation}
with reference pressure $p_o$~\cite{muller2010adaptive,giraldo2008study}.
For all remaining scenarions, we use the constants
\begin{equation}\label{eq:atmosphere-constants}
    \gamma = 1.4 ,\qquad \Pr =  0.71 ,\qquad R = 287.058 ,\qquad p_0 = \SI{10e5}{\Pa}, \qquad g = 9.8.
\end{equation}

We compute the initial background density and pressure by inserting the assumption of a constant background energy in \cref{eq:hydrostatic-balance}.
The background atmosphere is then perturbed.
\todo{What about our weird boundary condition for heat flux and energy?}
Let
\begin{equation}
  \label{eq:radius}
  r^2 = (x - x_0)^2 + (z - z_0)^2
\end{equation}
denote the difference between spatial positions $x,z$ and the center of a bubble $x_c, z_c$.
For our first scenario, the cosine bubble, we use a pertubation of the form
\begin{equation}
  \label{eq:cos-pertubation-alt}
  \pertubationPotT = 
    \nicefrac{A}{2} \left( 1 + \cos(\pi r) \right) \left[ r \leq a \right]
\end{equation}
\begin{align}
  \label{eq:cos-pertubation}
  \pertubationPotT &= \begin{cases}
    \nicefrac{A}{2} \left[ 1 + \cos(\pi r) \right] & r \leq a, \\
    0 & r > a,
    \end{cases}
\end{align}
where $A$ denotes the maximal perturbation and $a$ is the size of the bubble.
We use the constants
\begin{equation}\label{eq:cosine-bubble}
  A = 0.5 \quad a = \SI{250}{\m} \quad x_c = \SI{500}{\m} \quad z_c = \SI{350}{\m}.
\end{equation}
We use a constant viscosity of $\mu = 0.1$.
For our second scenario, the cold bubble on top of a warm bubble, we use the pertubations
\begin{equation}
  \label{eq:bubbles-pertubation}
  \pertubationPotT =
  \begin{cases}
    A & r \leq a, \\
    A \exp \left( - \frac{(r-a)^2}{s^2} \right) & r > a,
    \end{cases}
\end{equation}
where $s$ is the decay rate and $r$  is the radius to the center~\pcref{eq:radius}.
We have two bubbles, with constants
\begin{equation}
  \label{eq:bubbles-values}
\begin{alignedat}{6}
  & \text{warm:} \qquad && A = 0.5, \quad&& a = \SI{150}{\m}, \quad&& s = \SI{50}{\m}, \quad&& x_c = \SI{500}{\m,} \quad&& z_c = \SI{300}{\m},\\
  & \text{cold:} \qquad && A = -0.15, \quad&& a = \SI{0}{\m}, \quad&& s = \SI{50}{\m}, \quad&& x_c = \SI{560}{\m}, \quad&& z_c = \SI{640}{\m}.
  \end{alignedat}
\end{equation}
Similar to~\cite{muller2010adaptive} and to the previous scenario, we use a constant viscosity of $\mu = 0.1$ to regularize the solution.
To ensure that the atmosphere stays in hydrostatic balance~\cite{giraldo2008study}, we need to impose the viscous heat flux\todo{Show full bc?}
\begin{equation}
  \label{eq:atmosphere-bc}
  \viscFlux_{\QE} = \kappa \pdv{\overline{T}}{z}.% =
%\kappa \frac{R \overline{\theta} \left(\frac{\overline{p}{\left (z \right )}}{p_{0}}\right)^{\frac{R}{c_{p}}} \frac{d}{d z} \overline{p}{\left (z \right )}}{c_{p} \overline{p}{\left (z \right )}}.
\end{equation}
